\chapter{Base System Provisioning}\label{chap:base-provisioning}
\section{Hardware Requirements \& Baseline}
While the "Sovereign-Fort" architecture is designed to be resource-efficient, the use of ZFS and containerized workloads necessitates a specific hardware baseline to ensure data-integrity and system responsiveness.

\begin{itemize}
	\item \textbf{CPU :} Any x86\_64 processor with at least 4 cores (e.g., Intel Core i5-7400 or equivalent).
	\item \textbf{Memory :} Minimum 8GB RAM. (Note: ZFS utilizes adaptive replacement cache (ARC), so 16GB+ is recommended for heavy I/O workloads).
	\item \textbf{Storage :} 
	\begin{itemize}
		\item 1x SSD (min. 64GB) for the Debian 13 OS and Docker Root.
		\item 2x or more HDD/SSD for the ZFS \texttt{tank} pool (RAIDZ1 or Mirroring).
	\end{itemize}
	\item \textbf{GPU (Optional) :} Dedicated NVIDIA GPU (e.g., GTX 1060 6GB) for hardware-accelerated transcoding in services like Jellyfin or Immich.
\end{itemize}

\subsection{Test Environment}
The reference implementation for this manual utilizes a an \texttt{Intel i5-7400} CPU and a \texttt{GTX 1060} GPU with 16\unit{GB} of system RAM. The ZFS pool used is composed of three (3) 8\unit{TB} SeaGate BarraCuda HDDs in a Raid-Z1 configuration, for a total usable 14.4\unit{TB} storage pool.

\section{Prerequisites}


\begin{itemize}
	\item \textbf{Network :} An internet connection is needed for the installation, from start to finish and even during regular operations.
	\item \textbf{Installation Media :} An external storage of a minimum of 32\unit{GB} of storage is needed for the \texttt{netinst} ISO. A static IP is also necessary since manually configured IPs do not change automatically, potentially breaking your services.
	\item \textbf{External Gateway (Optional) :} Access to a VPS (Oracle \/ AWS \/ Azure \/ etc.) if following the Wireguard tunnel instructions in {Chapter }\ref{chap:security}
	
	
	
\end{itemize}