\chapter{Data Integrity \& Storage Hierarchy}\label{chap:storage}
\section{Mount Point Strategy}
To maintain a standardized environment, ZFS datasets are mounted directly to the Debian 13 filesystem using the \texttt{mountpoint} property. This ensures that storage is automatically managed by the ZFS kernel module rather than \texttt{/etc/fstab}, making the storage pool entirely portable.

\begin{itemize}
	\item \textbf{Host Mapping:} Datasets are mapped to the \texttt{/mnt} directory to separate infrastructure data from the OS root.
	\begin{itemize}
		\item \texttt{tank/appdata} $\rightarrow$ \texttt{/mnt/appdata}
		\item \texttt{tank/db} $\rightarrow$ \texttt{/mnt/db}
		\item \texttt{tank/media} $\rightarrow$ \texttt{/mnt/media}
	\end{itemize}
	\item \textbf{Container Mapping:} These standardized host paths are passed into Docker containers as bind-mounts (e.g., \texttt{/mnt/appdata/vaultwarden:/data}).
\end{itemize}


\begin{lstlisting}[language=bash, caption=Applying Professional Mount Points]
	# Set the mountpoint property to override the default
	/tank/ hierarchy
	zfs set mountpoint=/mnt/appdata tank/appdata
	zfs set mountpoint=/mnt/db tank/db
	zfs set mountpoint=/mnt/media tank/media
	
	# Verify the new hierarchy
	zfs list
\end{lstlisting}

\section{Essential ZFS Properties}
Beyond mount points and recordsizes, several pool-wide and dataset-specific properties are required to ensure the "Solid Foundation" of the server.

\begin{itemize}
	\item \textbf{Compression (lz4):} Enabled by default on all datasets. LZ4 is extremely fast and reduces the physical footprint of configuration files and databases without a noticeable CPU penalty.
	\item \textbf{Xattr (sa):} For Linux-based systems like Debian 13, setting \texttt{xattr=sa} stores extended attributes in the inode rather than hidden files, significantly improving performance for small file access.
	\item \textbf{Atime (off):} Disables access-time updates every time a file is read. This reduces unnecessary write operations, extending the lifespan of the HDDs/SSDs.
	\item \textbf{Canmount (off):} Set on the root \texttt{tank} dataset to prevent the system from mounting the parent pool directly, forcing strictly managed child dataset mounts.
\end{itemize}

\begin{lstlisting}[language=bash, caption=Optimizing Dataset Properties]
	# Apply performance optimizations
	zfs set compression=lz4 tank
	zfs set xattr=sa tank
	zfs set atime=off tank
	
	# Disable mounting for the root pool to keep /tank clean
	zfs set canmount=off tank
\end{lstlisting}

\section{Verification of Storage Logic}
A professional SOP requires verification steps to ensure the foundation is correctly laid before moving to automation.

\begin{lstlisting}[language=bash, caption=Verifying the Foundation]
	# Check that datasets are correctly hooked into /mnt
	findmnt | grep tank
	
	# Check the health and parity of the RAID-Z1 pool
	zpool status tank
\end{lstlisting}


\section{Cluster Storage}
There are two main ways to manage storage across a cluster, both of which have advantages and disadvantages.

\begin{itemize}
	\item \textbf{Distributed Storage} (Ceph / Longhorn) \\
	You can distribute storage across every machine, essentially syncing it and making it easily accessible. \\
	
	\begin{itemize}
		\item \textbf{Pros :} \\
		Extremely redundant storage. One node dies, the rest don't lose data at all.
		\item \textbf{Cons :} \\
		Large overhead in RAM and CPU. Worse for low-performance nodes in a cluster.
	\end{itemize}
	
	\item \textbf{Centralized Storage} \\
	You can have one node / server act as a central storage provider.
	
	\begin{itemize}
		\item \textbf{Pros :} \\
		Little to no overhead. Easy to backup the storage and provision the cluster.
		\item \textbf{Cons :} \\
		More vulnerable to hardware damage. If the main node dies, the remaining nodes lose all data. 
	\end{itemize}
\end{itemize}

In my environment, I have one server acting as centralized storage, providing volumes to services and node. 

Distributed storage will be revisited later in the future. For now, I will stick to what I do know : ZFS.